---
title: "assignment1"
author: Yixiang Xu 36383724
format: 
  html:
   toc: true
  pdf:
   toc: true
execute:
  warning: false
  message: false
  echo: false
---

```{r}
library(tidyverse)
data_first<-read.csv("data/36383724_Yixiang Xu.csv")
```

## Introduction
This report investigates whether higher marketing spend per customer leads to higher purchase amounts. Understanding this relationship is important for optimizing marketing budget allocation. The analysis uses customer-level campaign data. The goal is to determine if increased spending improves revenue outcomes.

## Research Question
Does higher marketing spend per customer lead to higher purchase amounts?  
This report explores the relationship between marketing cost and customer spending behaviour.

## Dataset Introduction
The dataset contains information on individual customers, including marketing spend and purchase amounts.  
Data cleaning involved removing missing values and ensuring correct data types.  
Outliers were inspected and handled where necessary.  
The dataset is treated as the canonical record of the campaign.  
All preprocessing steps were conducted before analysis.

During this data preprocessing phase, we performed an in-depth cleaning of the raw dataset using the R language’s tidyverse toolchain. First, we used the `parse_date_time` function to perform a robust parsing of the `signup_date` column using various date formats (ymd, mdy, dmy), thereby resolving issues with inconsistent date formats. Next, for non-standard placeholders in the `notes` column, we used `na_if` to explicitly map empty strings and “N/A” to native missing values (NA), ensuring statistical consistency. To accurately assess actual revenue, we performed business logic adjustments, setting the transaction amount to 0 for all entries marked as “returned.” Finally, we cast `visits` to an integer type and `segment` to a categorical type, thereby constructing a clean dataset with standardized formats, rigorous logic, and suitability for subsequent statistical modeling.
```{r}
cleaned_data <- data_first %>%
  mutate(signup_date = parse_date_time(signup_date, orders = c("ymd", "mdy","dmy"))) %>%
  
  mutate(notes = na_if(notes, ""),
  notes = na_if(notes, "N/A")) %>%
  
  mutate(purchase_amount = if_else(notes == "returned" & !is.na(notes), 0, purchase_amount)) %>%

  mutate(
    visits = as.integer(visits),
    segment = as.factor(segment)
  )
```