---
title: "assignment1"
author: Yixiang Xu 36383724
format: 
  html:
   toc: true
  pdf:
   toc: true
execute:
  warning: false
  message: false
  echo: false
---

```{r}
library(tidyverse)
data_first<-read.csv("data/36383724_Yixiang Xu.csv")
```

# Introduction
This report examines whether an increase in marketing spend per customer leads to higher purchase amounts. Understanding this relationship is critical for optimizing the allocation of marketing budgets. This analysis uses customer-level marketing campaign data to determine whether increased marketing spend can improve revenue outcomes.

# Research Question
This report examines whether higher marketing spending by customers leads to higher purchase amounts and investigates the relationship between marketing costs and customer spending behavior.

# Dataset Introduction
The dataset contains information on individual customers, including marketing spend and purchase amounts.  
Data cleaning involved removing missing values and ensuring correct data types.  
Outliers were inspected and handled where necessary.  
The dataset is treated as the canonical record of the campaign.  
All preprocessing steps were conducted before analysis.
**data cleaning**
During this data preprocessing phase, we performed an in-depth cleaning of the raw dataset using the R language’s tidyverse toolchain. First, we used the `parse_date_time` function to perform a robust parsing of the `signup_date` column using various date formats (ymd, mdy, dmy), thereby resolving issues with inconsistent date formats. Next, for non-standard placeholders in the `notes` column, we used `na_if` to explicitly map empty strings and “N/A” to native missing values (NA), ensuring statistical consistency. To accurately assess actual revenue, we performed business logic adjustments, setting the transaction amount to 0 for all entries marked as “returned.” Finally, we cast `visits` to an integer type and `segment` to a categorical type, thereby constructing a clean dataset with standardized formats, rigorous logic, and suitability for subsequent statistical modeling.

## Dataset Description
The dataset used in this analysis provides a granular look at customer acquisition and spending patterns.It consists of `r ncol(cleaned_data)` variables and `r nrow(cleaned_data)` observations.The variables include a mix of character strings (customer IDs), date-time objects (signup dates), numeric doubles (marketing spend and purchase amounts), and integers (visit frequency). This variety allows for a multi-dimensional assessment of how campaign investments translate into revenue.
```{r}
cleaned_data <- data_first %>%
  mutate(signup_date = parse_date_time(signup_date, orders = c("ymd", "mdy","dmy"))) %>%
  
  mutate(notes = na_if(notes, ""),
  notes = na_if(notes, "N/A")) %>%
  
  mutate(purchase_amount = if_else(notes == "returned" & !is.na(notes), 0, purchase_amount)) %>%

  mutate(
    visits = as.integer(visits),
    segment = as.factor(segment)
  )
```

@tbl-summary aggregates purchase amounts for each customer segment. First, it groups the data based on the `segment` variable, then calculates several statistical measures, including the mean, median, minimum, and maximum of the purchase amounts. During the calculation, the effect of missing values is accounted for by setting `na.rm = TRUE`.
```{r}
#| label: tbl-summary
#| tbl-cap: summary segment
purchase_summary <- cleaned_data %>%
  group_by(segment)%>%
  summarise(
    Mean = mean(purchase_amount, na.rm = TRUE),
    Median = median(purchase_amount, na.rm = TRUE),
    Min = min(purchase_amount, na.rm = TRUE),
    Max = max(purchase_amount, na.rm = TRUE)
  )
```

@fig-summary are used to visualize the distribution of purchase amounts across different customer segments. The `ggplot()` function is used to specify the dataset, mapping `segment` to the x-axis and `purchase_amount` to the y-axis. The `geom_boxplot()` function generates the box plot to display the range and central tendency of purchase amounts within each customer segment. The plot is filled with sky blue, and titles and axis labels are added using `labs()` to make the chart clearer and easier to read.
```{r}
#| label: fig-summary
#| fig-cap: summary segment
ggplot(cleaned_data, aes(x = segment, y = purchase_amount)) +
  geom_boxplot(fill = "skyblue") +
  labs(
    title = "Purchase Amount Distribution by Customer Segment",
    x = "Customer Segment",
    y = "Purchase Amount"
  )
```

# result
@fig-linear illustrates the relationship between marketing spend and purchase amount. Each point in the scatter plot represents an observation in the dataset, and the red line indicates the fitted linear regression trend. The figure shows an overall positive correlation, meaning that as marketing spend increases, purchase amounts also tend to rise. This suggests that marketing investment may influence consumer purchasing behavior. This analysis is based on `nrow(cleaned_data)` observations.
```{r}
#| label: fig-linear
#| fig-cap: Relationship between marketing spend and purchase amount.
ggplot(cleaned_data, aes(x = marketing_spend, y = purchase_amount)) +
  geom_point(color = "steelblue", alpha = 0.6) +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  labs(
    title = "Relationship between Marketing Spend and Purchase Amount",
    x = "Marketing Spend",
    y = "Purchase Amount"
  )
```

@fig-segment illustrates the relationship between marketing spend and purchase amount across different customer segments. Each data point in the figure represents an observation, and the trend lines of different colors indicate the linear relationship for each customer group. The figure shows that, across all customer segments, there is an overall positive correlation between marketing spend and purchase amount; that is, as marketing spend increases, purchase amount typically increases as well. Although the overall trend is similar, the slopes of the trend lines vary slightly across different customer segments, suggesting that the strength of this relationship may differ among them. These results further support the research hypothesis that there is an association between marketing investment and higher purchase amounts.
```{r}
#| label: fig-segment
#| fig-cap: Relationship between marketing spend and purchase amount across customer segments.
ggplot(cleaned_data, aes(x = marketing_spend, y = purchase_amount, color = segment)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Marketing Spend vs Purchase Amount by Customer Segment",
    x = "Marketing Spend",
    y = "Purchase Amount",
    color = "Customer Segment"
  )
```

